FPGA Based Modelling of an ADPLL Network\vspace{-0.3cm}

### 2019-10-14 conf_paper/paper.tex
supposedly chanaged

\title{FPGA Based Modelling of an ADPLL Network\\}

\author{\IEEEauthorblockN{C. Dooley$^1$, E. Blokhina$^1$, B. Mulkeen$^1$, D. Galayko$^2$}
    \IEEEauthorblockA{\textit{$^1$School of Electrical \& Electronic Engineering, University College Dublin}\\
    \textit{$^2$LIP6 Lab, Sorbonne Université}
    	\vspace{-0.8cm}
	}
}
\maketitle

@@

    This paper introduces and compares the implementation of a number of FPGA based ADPLL network prototyping architectures. Comparison is then made between the expected performance and role of each method as a development tool.

@@

The technical motivation underlying this paper is the creation of an Field Programmable Gate Array (FPGA) based prototyping platform for All Digital Phase Lock Loop (ADPLL) networks. As the creation of Application Specific Integrated Circuits (ASICs) is an expensive and time consuming process, and mistakes made in the design may require the manufacture of a new chip, it is important to ensure that any errors made in the design have been located and addressed before manufacture..

@@

The particular system that this paper will discuss is an ADPLL network. This is based on an idea first proposed in a 1995 paper by Pratt and Nguyen \cite{pratt1995distributed}, in which they proposed an alternative clock distribution network for ICs using a Cartesian grid of clocking areas, each with their own Phase Lock Loop, which has become known as a PLL Network . This methodology was implemented by Gutnik et. al \cite{gutnik2000active} who found it to be feasible. Subsequently Javiden et. al
\cite{javidan2011all} proposed the implementation of such a system using All Digital PLLs, which better suited the use cases of the technology and avoided some of the flaws pointed out in the original paper. Zianbetov and Shan followed on from this proposition by implementing such a network on a custom IC \cite{zianbetov2013phd,shan2014phd}.\\

@@

\section{System Architecture}
An ADPLL network, as the name would suggest, is created from a number of ADPLLs that are coupled using digital phase comparators which attempt to measure the phase error between two oscillators, with each non edge node being connected to four neighbours. Figure \ref{fig:adpll_network} illustrates such a system on a small scale, with a number of ADPLLs laid out in a Cartesian grid. The individual ADPLLs, often called distributed ADPLLs as the phase detectors are shared between nodes, are made up of a digitally controllable oscillator, an error combination block and a digital loop filter with the aforementioned digital phase detectors lying between each pair of oscillators.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.25\textwidth]{adpll_network}
    \caption{ADPLL Network Architecture.}
    \label{fig:adpll_network}
    \vspace{0.5cm}
    \includegraphics[width=0.4\textwidth]{dist_adpll}
    \caption{Distributed ADPLL Design.}
    \label{fig:adpll_base}
    \vspace{-0.5cm}
\end{figure}
\subsection{Digitally Controlled Oscillators}
While regular Voltage Controlled Oscillator (VCO) is tuned by the input voltage over a continuous range of values, in a digital system there are a very limited number of voltages representable, most commonly just two levels. As such, using the voltage itself to control an oscillator is not a viable strategy and instead Digitally (or Numerically) Controlled Oscillators (DCO) accept an \textit{n} bit wide control signal that alters the period of the output waveform. While tuning range, centre frequency and linearity apply as in a VCO, a DCO also has a minimum frequency step as a result of the non continuous control values. %TODO WORDING
In combination with the bit width of the control signal, the minimum frequency step limits the range over which the oscillator can be tuned. Frequently the output of the DCO is divided down to allow the control blocks in the feedback path that are clocked on this signal, to run at a lower and thus easier frequency.
\subsection{Loop Filter}
The loop filter is a key aspect of a phase lock loop as without it only the current output of the phase detector could be used to compute the control voltage of the VCO.
In an ADPLL it performs an identical role however, the digital environment results in the traditional loop filter being replaced by a discrete time Proportional Integral (PI) controller, with the integral path of the controller implemented by an accumulator.
\subsection{Error Combiner}
The Error Combiner, as the name suggests, performs a summation of the error signals from the phase detectors lying between any neighbouring oscillator and its neighbours. It is often convenient for this to be performed as a weighted sum, thus avoiding different gain requirements for the loop filter of oscillators in the corners of the grid that only have a pair of neighbours. The ability to change the weights during operation enables ``uni-directional start-up'', which Javiden et. al \cite{javidan2011all} found prevented the system from entering unwanted stable equilibriums warned of in Pratt and Nguyen's seminal paper.
\subsection{Phase Detector}
In a conventional PLL the phase detector measures a value proportional to the difference in phase between a pair of signals, which is then output as a continuous voltage.
A digital phase detector then attempts to do the same by measuring the time difference between fixed reference points in the waveform, which as a digital system deals with square waves, is typically the rising edge of each input to the block.
As in the oscillator, quantisation also plays a role in the behaviour of this block, with the time difference between each rising edge having a minimum possible measurement step.
In distributed ADPLLs the phase detector has dual outputs, each the inverse of the other, in order to feed the correct value to both of the connected error combiners.

@@

            \end{tabular}
        \end{tiny}
        \caption{}
        \label{table:b_c_perf}
    \end{center}
\vspace{-0.5cm}
\end{table}
From the table it is apparent that the coarser period and detection steps of Design ``1A'' (3.875 nanoseconds as at greater clock frequencies timing violations would occur) cannot deliver the same jitter performance as the inverter ring driven designs. Looking at the two subsequent designs their lesser flexibility is made up for by their more realistic jitter behaviour. As would be expected best performance occurs when the network is in PLL mode and as the quality of the reference degrades as the network is connected up performance becomes worse. In terms of cycle-to-cycle jitter the best performing design is that of ``2A'', however, its time interval error is significantly worse especially in bilateral network mode. This more aggressive degradation as the network is linked up can be attributed to the worse detection resolution of the FPGA clocked phase detector, four times worse than the 1.2 nanosecond step of the entirely inverter based Design ``2B''.\\
The further a PLL is away from the reference feeding ADPLL 11 the worse performance is, peaking furthest from the reference in ADPLL 22.
The lesser cycle-to-cycle of Design ``2A'' follows from the implementation independent design of the phase detector. Unlike the inverter based design where the implementation may result in the propagation delay between taps on the delay line changing, from both tap to tap in the same detector and the same tap in different detectors, the clocked design has identical detection resolutions as each register is driven by the FPGA clock. This can be highlighted in the more easily demonstrable case of the clocked and inverter based oscillators.\\
\begin{figure}[h]%
    \centering
    \subfloat[Inverter Based Design.]{{\includegraphics[width=0.5\linewidth]{distrib_ring} }}%
    \subfloat[FPGA Clocked Design.]{{\includegraphics[width=0.5\linewidth]{distrib_pa} }}%
    \caption{Example Distribution of Periods.\label{fig:dists}}    
    \vspace{-0.5cm}
\end{figure}
Figure \ref{fig:dists} paints a picture of the unrealistic nature of the FPGA clocked design in comparison to its inverter based counterpart, each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment. This effects the suitability of the FPGA clocked designs for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network.
\section*{Conclusion}
An FPGA is a suitable environment for the prototyping of ADPLL networks and can fill a number of roles depending on the designer's requirements. However, the designer should bear in mind the platform's limitations, especially when attempting to use a frequency that is not orders of magnitude lower than the maximum clock available from the FPGA's clock manager. Inverter based and FPGA clocked designs, both have a niche in which they are the preferred option, with the clocked design better suiting replication of an existing design at scaled-down frequencies and the inverter based designs better in the modelling and analysis of ADPLL network behaviour.

>>>>

\title{FPGA Based Modelling of an ADPLL Network\vspace{-0.3cm}}
\author{
	\IEEEauthorblockN{C. Dooley$^1$, E. Blokhina$^1$, B. Mulkeen$^1$, D. Galayko$^2$}
    \IEEEauthorblockA{\textit{$^1$School of Electrical \& Electronic Engineering, University College Dublin}\\
    \textit{$^2$LIP6 Lab, Sorbonne Universit\'e}
    	\vspace{-0.7cm}
	}
}
\IEEEoverridecommandlockouts
\IEEEpubid{\makebox[\columnwidth]{978-1-7281-1201-5/19/\$31.00~\copyright2019 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}
\maketitle

@@

    This paper introduces and compares the implementation of a number of FPGA based ADPLL network prototyping architectures. Networks are then created using three different ADPLL implementations and tests performed on each. Based on these test results, comparison is made to both the expected performance and role of each ADPLL design as a development tool.

@@

The technical motivation underlying this paper is the creation of an Field Programmable Gate Array (FPGA) based prototyping platform for All Digital Phase Lock Loop (ADPLL) networks. As the creation of Application Specific Integrated Circuits (ASICs) is an expensive and time consuming process, and mistakes made in the design may require the manufacture of a new chip, it is important to ensure that any errors made in the design have been located and addressed before manufacture.

@@

The particular system that this paper will discuss is an ADPLL network. This is based on an idea first proposed in a 1995 paper by Pratt and Nguyen \cite{pratt1995distributed}, in which they proposed an alternative clock distribution network for ICs using a Cartesian grid of clocking areas, each with their own Phase Lock Loop, which has become known as a PLL Network . This methodology was implemented by Gutnik et. al \cite{gutnik2000active} who found it to be feasible. Subsequently Javiden et. al \cite{javidan2011all} proposed the implementation of such a system using All Digital PLLs, which better suited the use cases of the technology and avoided some of the flaws pointed out in the original paper. Zianbetov and Shan followed on from this proposition by implementing such a network on a custom IC \cite{zianbetov2013phd,shan2014phd}.\\

@@

\section{System Architecture}
An ADPLL network, as the name would suggest, is created from a number of ADPLLs that are coupled using digital phase comparators which attempt to measure the phase error between two oscillators, with each non edge node being connected to four neighbours. Figure \ref{fig:adpll_network} illustrates such a system on a small scale, with a number of ADPLLs laid out in a Cartesian grid. The individual ADPLLs, often called distributed ADPLLs as the phase detectors are shared between nodes, are made up of a digitally controllable oscillator, an error combination block and a digital loop filter with the aforementioned digital phase detectors lying between each pair of oscillators.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.25\textwidth]{adpll_network}
    \caption{ADPLL Network Architecture.}
    \label{fig:adpll_network}
    \vspace{0.5cm}
    \includegraphics[width=0.4\textwidth]{dist_adpll}
    \caption{Distributed ADPLL Design.}
    \label{fig:adpll_base}
    \vspace{-0.5cm}
\end{figure}
\subsection{Digitally Controlled Oscillators}
While regular Voltage Controlled Oscillator (VCO) is tuned by the input voltage over a continuous range of values, in a digital system there are a very limited number of voltages representable, most commonly just two levels. As such, using the voltage itself to control an oscillator is not a viable strategy and instead Digitally (or Numerically) Controlled Oscillators (DCO) accept an \textit{n} bit wide control signal that alters the period of the output waveform. While tuning range, centre frequency and linearity apply as in a VCO, a DCO also has a minimum frequency step as a result of the non continuous control values. %TODO WORDING
In combination with the bit width of the control signal, the minimum frequency step limits the range over which the oscillator can be tuned. Frequently the output of the DCO is divided down to allow the control blocks in the feedback path that are clocked on this signal, to run at a lower and thus easier frequency.\vspace{-0.2cm}
\subsection{ Loop Filter}
The loop filter is a key aspect of a phase lock loop as without it only the current output of the phase detector could be used to compute the control voltage of the VCO.
In an ADPLL it performs an identical role however, the digital environment results in the traditional loop filter being replaced by a discrete time Proportional Integral (PI) controller, with the integral path of the controller implemented by an accumulator.\vspace{-0.2cm}
\subsection{Error Combiner}
The Error Combiner, as the name suggests, performs a summation of the error signals from the phase detectors lying between any neighbouring oscillator and its neighbours. It is often convenient for this to be performed as a weighted sum, thus avoiding different gain requirements for the loop filter of oscillators in the corners of the grid that only have a pair of neighbours. The ability to change the weights during operation enables ``uni-directional start-up'', which Javiden et. al \cite{javidan2011all} found prevented the system from entering unwanted stable equilibriums warned of in Pratt and Nguyen's seminal paper.\vspace{-0.2cm}
\subsection{Phase Detector}
In a conventional PLL the phase detector measures a value proportional to the difference in phase between a pair of signals, which is then output as a continuous voltage.
A digital phase detector then attempts to do the same by measuring the time difference between fixed reference points in the waveform, which as a digital system deals with square waves, is typically the rising edge of each input to the block.
As in the oscillator, quantisation also plays a role in the behaviour of this block, with the time difference between each rising edge having a minimum possible measurement step.
In distributed ADPLLs the phase detector has dual outputs, each the inverse of the other, in order to feed the correct value to both of the connected error combiners.

@@

            \end{tabular}
        \end{tiny}
        \caption{ADPLL Performance Comparison.}
        \label{table:b_c_perf}
    \end{center}
\vspace{-0.5cm}
\end{table}
From the table it is apparent that the coarser period and detection steps of Design ``1A'' (3.875 nanoseconds as at greater clock frequencies timing violations would occur) cannot deliver the same jitter performance as the inverter ring driven designs. Looking at the two subsequent designs their lesser flexibility is made up for by their more realistic jitter behaviour. As would be expected best performance occurs when the network is in PLL mode and as the quality of the reference degrades as the network is connected up performance becomes worse. In terms of cycle-to-cycle jitter the best performing design is that of ``2A'', however, its time interval error is significantly worse especially in bilateral network mode. This more aggressive degradation as the network is linked up can be attributed to the worse detection resolution of the FPGA clocked phase detector, four times worse than the 1.2 nanosecond step of the entirely inverter based Design ``2B''.\\
The further a PLL is away from the reference feeding ADPLL 11 the worse performance is, peaking furthest from the reference in ADPLL 22.
The lesser cycle-to-cycle of Design ``2A'' follows from the implementation independent design of the phase detector. Unlike the inverter based design where the implementation may result in the propagation delay between taps on the delay line changing, from both tap to tap in the same detector and the same tap in different detectors, the clocked design has identical detection resolutions as each register is driven by the FPGA clock. This can be highlighted in the more easily demonstrable case of the clocked and inverter based oscillators.\\
\begin{figure}[h]%
	\vspace{-0.8cm}
    \centering
    \subfloat[Inverter Based Design.]{{\includegraphics[width=0.5\linewidth]{distrib_ring} }}%
    \subfloat[FPGA Clocked Design.]{{\includegraphics[width=0.5\linewidth]{distrib_pa} }}%
    \caption{Example Distribution of Periods.\label{fig:dists}}    
    \vspace{-0.4cm}
\end{figure}
Figure \ref{fig:dists} paints a picture of the unrealistic nature of the FPGA clocked design in comparison to its inverter based counterpart, each oscillator will have identical intrinsic period steps thus eliminating any jitter that maybe be seen in the system due to misalignment. This effects the suitability of the FPGA clocked designs for the improvement of simulation models, or the examination of how a novel block may impact the behaviour of the network.
\section*{Conclusion}
An FPGA is a suitable environment for the prototyping of ADPLL networks and can fill a number of roles depending on the designer's requirements. However, the designer should bear in mind the platform's limitations, especially when attempting to use a frequency that is not orders of magnitude lower than the maximum clock available from the FPGA's clock manager. Inverter based and FPGA clocked designs, both have a niche in which they are the preferred option, with the clocked design better suiting replication of an existing design at scaled-down frequencies and the inverter based designs better in the modelling and analysis of ADPLL network behaviour.
